---
title: "Tree-based Method"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8,fig.height=6, warning = FALSE, fig.align='center')
library(tree)
library(randomForest)
library(MASS)
library(RColorBrewer)
library(plotly)
library(ggplot2)
library(ISLR)
library(gbm)
library(class)
```

```{r}
data <- read.csv("~/Documents/BU/2016 fall/MA 675 STAT PRAC 1/consulting/Dental & Age/data/Faillace - for Stats Consulting.csv")
teeth <- data[c(1:583),]
teeth <- teeth[-c( 427, 143, 78, 236, 307, 363, 373, 114, 361, 144, 131, 95, 13, 126, 22, 9, 124, 461, 7, 6, 438, 344, 5, 107, 226, 199, 343, 240, 531, 244),-c(1)]
teeth <-subset(teeth, !is.na(teeth$Age))
```

```{r}
teeth1=na.omit(teeth)
```


```{r}
set.seed(1)
train = sample(1:nrow(teeth1), nrow(teeth1)/2)
ttrain = teeth1[train,]
ttest = teeth1[-train,]
```

Tree based method is attractive because of the ease of interpretability.

### tree & prune tree
```{r, fig.width=8,fig.height=8}
tree.teeth = tree(Age~., ttrain)
plot(tree.teeth)
text(tree.teeth,pretty=0)
mean((predict(tree.teeth,ttest)-ttest$Age)^2)

set.seed(11)
cv.teeth = cv.tree(tree.teeth, FUN = prune.tree)
plot(cv.teeth$size,cv.teeth$dev,type='b')

prune.teeth = prune.tree(tree.teeth,best=10)
# plot(prune.Carseats)
# text(prune.Carseats,pretty=0)
mean((predict(prune.teeth,ttest)-ttest$Age)^2)
plot(prune.teeth)
text(prune.teeth,pretty=0)
```


### bagging & random forest
```{r}
xtrain = ttrain[, -1]
xtest = ttest[, -1]
ytrain = ttrain[, 1]
ytest = ttest[, 1]

testerrors=data.frame("ntree"=rep(1:500,28), 
                      "mtry"=rep(c(1:28), each=500),
                      "TestError"=NA)

for (i in 1:28) {
  m = randomForest(xtrain, ytrain, xtest = xtest, ytest = ytest, 
    mtry = i, ntree = 500)
  testerrors[(i*500-499):(i*500),3]=m$test$mse
}

testerrors$mtry <- as.factor(testerrors$mtry)
```

```{r}
p <- ggplot(data = testerrors,
       aes(x = ntree, y = TestError, group = mtry, colour = mtry)) +
  geom_line(size=0.8) + 
  theme(plot.title = element_text(size=15, face="bold", vjust=2),
        axis.text.x=element_text(size=10, vjust=0.35)) +
  labs(title = "Test Errors for Random Forests") + 
  xlab("Number of Trees") + ylab("Test Errors") +
  scale_colour_brewer(palette = "Paired")
gg <- ggplotly(p) 
layout(gg)
```

* 96/97 a little bit smaller than tree and prune tree


### Generalized Boosted Regression 
However, the trees are not the best model in terms of prediction.  Improved version of the tree gives more accurate prediction with the cost of loosing clear interpretation.

```{r}
boost.teeth=gbm(Age~.,
                 data=teeth1,
                 distribution="gaussian",
                 n.trees=5000,
                 interaction.depth=4,
                 shrinkage=0.2,
                 cv.folds=10,
                 verbose=F)
```

```{r}
gbm.perf(boost.teeth, method="OOB")

# right?
boost.pred = predict(boost.teeth, ttest, n.trees=66)
mean((boost.pred-ttest$Age)^2)
```


```{r}
best.iter <- gbm.perf(boost.teeth,method="OOB",plot.it=FALSE)
print(best.iter)
summary(boost.teeth,n.trees=best.iter) # based on the estimated best number of trees

best.iter <- gbm.perf(boost.teeth,method="cv",plot.it=FALSE)
print(best.iter)
summary(boost.teeth,n.trees=best.iter) # based on the estimated best number of trees

plot.gbm.tree<-function(object,n.trees = NULL) {
  summary.object<-summary(object,plotit=FALSE,n.trees=n.trees)
  reorder_size <- function(x) {
    factor(x, levels = names(sort(table(x))))
  }
  dttmp<-summary.object
  dttmp$variables <- factor(dttmp$var, 
                            levels = dttmp$var[order(dttmp$rel.inf)])
  colourCount = length(dttmp$variables)
  color_levels<-colorRampPalette(brewer.pal(9, "Blues"))(colourCount)
  gp<- ggplot(dttmp) + geom_bar() + 
  aes( x=variables, weight = rel.inf , fill=variables)+
  ylab("Relative influence")+ coord_flip()+
   scale_fill_manual(values = color_levels )+ theme(legend.position="none")
  print(gp)
}
plot.gbm.tree(boost.teeth, best.iter) 
yhat.boost=predict(boost.teeth,newdata=teeth1,n.trees=best.iter) 
plot(teeth1$Age,yhat.boost)
```


```{r,fig.width=8,fig.height=8}
par(mfrow=c(2,2))
plot(boost.teeth,i="X24")
plot(boost.teeth,i="X9")
plot(boost.teeth,i="Yun.31")
plot(boost.teeth,i="X11")
```


